---
title: "SCO_Project1_Group1"
author: "Brian Lakey, Reshma Sekar, Julia Wu, Carlton Washburn"
date: "February 10, 2016"
output: html_document
---




##**PROBLEM 1**
####Formulate the dedicated portfolio construction problem as a linear program  

#### Setting up parameters for example portfolio  

decision variables: (x1), ... , (x10) amount of each bond  
                    (z1), ... , (z7) excess cash at the end of each year (excluding the excess cash at the                                         last period)  

minimize: 102(x1) + 99(x2) + 101(x3) + 98(x4) + 98(x5) + 104(x6) + 100(x7) + 101(x8) + 102(x9) + 94(x10)
 cost of each bond

subject to:  
 year 1: 12000 = 100(x1) + 5(x1) + 3.5(x2) + 5(x3) + 3.5(x4) + 4(x5) + 9(x6) + 6(x7) + 8(x8) + 9(x9) + 7(x10) - (z1)  
 year 2: ...  

constraints = years (8)  
variables = bond terms (10), excess cash (7)  

####Setting up parameters for example portfolio
```{r}
library('lpSolveAPI')

p = c(102,99,101,98,98,104,100,101,102,94) #price
c = c(5,3.5,5,3.5,4,9,6,8,9,7) #coupon
m = c(1,2,2,3,4,5,5,6,7,8) #maturity
l = c(12000,18000,20000,20000,16000,15000,12000,10000) #liabilities (RHS)
```


####Setting up code for solving optimization
```{r}
my.lp = make.lp(length(l),length(p)+length(l)-1) #constraints, variables

i=1
while (i <= length(m)) { # cycles through the bonds
  v = c(rep(0,length(l))) # assigns all variables 0s
  v[1:m[i]] = c[i] # assigns coupons for all eligible years
  v[m[i]] = c[i]+100 # adds payout for final year
  set.column(my.lp,i,v) # writes to columns
  i = i+1 # increments
}

k=1
while (k <= (length(l)-1)) { # cycles through years
  z = c(rep(0,length(l))) # assigns all excess cash variables to 0
  z[k] = -1 # assigns -1 to current year excess cash
  if (k < length(l)) { 
    z[k+1] = 1 # assigns 1 to previous year excess cash
  }
  set.column(my.lp,k+length(m),z)
  k = k+1 #increments
}

set.objfn(my.lp, c(p, rep(0,(length(l)-1))))
set.constr.type(my.lp, rep(">=",length(l)))
set.rhs(my.lp, l)
```

  
##**PROBLEM 2**
####Solve the example portfolio case in R using lpSolveAPI() instead lpSolve()  
  
####Solving
```{r}
solve(my.lp)
```
  
####Getting optimal variables
```{r}
get.variables(my.lp)
```

    
##**PROBLEM 3**
#### Write a function in R that can construct a portfolio for any set of liabilities and bonds  

####Function takes in bond portfolio parameters and return a .lp object
```{r}
dedicate_g1 = function (p, c, m, l) {
  my.lp = make.lp(length(l),length(p)+length(l)-1)
  
  i=1
  while (i <= length(m)) {
    v = c(rep(0,length(l)))
    v[1:m[i]] = c[i]
    v[m[i]] = c[i]+100
    set.column(my.lp,i,v)
    i = i+1
  }
  
  k=1
  while (k <= (length(l)-1)) {
    z = c(rep(0,length(l)))
    z[k] = -1
    if (k < length(l)) {
      z[k+1] = 1
    }
    set.column(my.lp,k+length(m),z)
    k = k+1
  }
  
  set.objfn(my.lp, c(p, rep(0,(length(l)-1))))
  set.constr.type(my.lp, rep(">=",length(l)))
  set.rhs(my.lp, l)
  
  return(my.lp)
}
```



####Testing function on example portfolio
```{r}
old_port = dedicate_g1(p,c,m,l)
solve(old_port)
```

```{r}
get.variables(old_port)
#looks right
```

```{r}
get.objective(old_port)
#looks right
```

  
##**PROBLEM 4**
####Construct a dedicated portfolio using this liability stream and current bond information  

```{r}
setwd("C:/Users/brian/Desktop/MS-BA/03 Spring Semester/Stochastic Control and Optimization/sco_project 1")
bonds = read.csv('treasury_bonds.csv', header=TRUE) # we creted a csv document containing all of the
                  # important bond fields from the provided website
#names(bonds)
```

  
####Creating a column of date objects
```{r}
bonds['Matur_Date'] = NaN
bonds$Matur_Date = as.Date(as.character(bonds$Maturity),format="%m/%d/%Y")
```


####Creating a list of liability dates
```{r}
liability_dates = c('6/30/2016','12/31/2016','6/30/2017','12/31/2017','6/30/2018','12/31/2018','6/30/2019',
                    '12/31/2019','6/30/2020','12/31/2020','6/30/2021','12/31/2021')
liability_dates = as.Date(liability_dates,format="%m/%d/%Y")
```

####Adding portfolio start date
Note: we set out portfolio start date a 12/31/2015, six months prior to the first liability date
```{r}
start_date = as.Date("12/31/2015", format="%m/%d/%Y")
portfolio_dates = c(start_date,liability_dates)
```


####Assigning maturity periods using the date lists
Goal is to have list of maturity due periods (numbering 1-12), similar to format given in example portfolio
```{r}
bonds['period'] = NaN
for (i in c(1:nrow(bonds))) {
  k=1
  while (k <= length(portfolio_dates)) {
    if (bonds$Matur_Date[i] < portfolio_dates[k]) {
      bonds$period[i] = k-1
      k = length(portfolio_dates)+1
    } else {
      k = k+1
    }
  }
}
```


####Trim all bonds maturing after last liability date
```{r}
w = sum(!is.na(bonds$period)) # find number of bonds that mature during project 
```

####Using function to optimize new portfolio
```{r}
pp = bonds$Asked[1:w] #price, removing all bonds that do not mature during the life of the project
cc = bonds$Coupon[1:w] #coupon
mm = bonds$period[1:w] #maturity
ll = c(9000000,9000000,10000000,10000000,6000000,6000000,9000000,9000000,10000000,10000000,5000000,3000000) #liabilities (RHS)

new_port = dedicate_g1(pp,cc,mm,ll)
#sum(l) #pv of liabilities = $96,000,000
#optimal portfolio = $74,042,228

solve(new_port)
```

Last 11 variables are zs - excess cash carried over
```{r}
get.variables(new_port)
```

Savings of $21,957,772 (ignoring interest/time value of money)
```{r}
get.objective(new_port)
```

####Plotting bond and liability duals versus libability period (analagous to date)
```{r}

duals = get.dual.solution(new_port)
sensitivity = get.sensitivity.rhs(new_port)
dual.df = data.frame(sensitivity$duals,sensitivity$dualsfrom,sensitivity$dualstill)

liab_duals = dual.df[1:12,]
bond_duals = dual.df[13:235,]

```

The graph below shows the shadow prices for each of the 224 bonds by maturity period. The shadow price represents the increase in the objective value from an additional purchase of a bond. We see a positive linear relationship between maturity and the duals, which makes sense because the further into the future a bond matures, the more coupons it will generate and the more expensive it is. For example, bond #1 has a maturity of 2/15/2016 and a shadow price of 8.05. This means buying an additional unit of bond 1 would increase the objective value by 8.05.
```{r}

plot(mm, bond_duals$sensitivity.duals, main="Bond Duals by Maturity Period")

```

Looking at liability duals by maturity shows a negative linear relationship. The first liability has a shadow price of 0.92 which means that an increase in the first constraint by $1 requires a $0.92 increase in the objective value. Shadow prices fall as maturity dates rise because there is more time to accumulate return to cover the liability.
```{r}

plot(c(1:12), liab_duals$sensitivity.duals, main="Liability Duals by Maturity Period")

```




